---
title: "Typical .CEL file preprocessing in R"
output: html_notebook
---

For this test, we'll use any of the .CEL files published in [this](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?token=trmzbecoaqyugtc&acc=GSE43358) study. We'll then extract the .CEL file and proceed to preprocess it using R.

The code blocs are executed by using *Ctrl+Shift+Enter*. 

We name the file *tesq.CEL* and put it in the same directory as the notebook.

```{r}
library("affy")

cel = dir(path="tesq.CEL")
samples = ReadAffy(filenames=cel)
arr = rma(samples)
expressions = exprs(arr)
print(expressions)
```

The array *expressions* is an array of [gene expression](https://en.wikipedia.org/wiki/Gene_expression) values in breast tissue (normalized according to the documentation). The base hypothesis is that we can use those values as low level features a neural network could exploit to make predictions on phenotypic traits(in our case, the probability of breast cancer occurring). We persist the array on disk as a .csv file to pursue  further processing in Python:

```{r}
write.table(expressions, file="expressions.csv", quote=F, col.names=c("id;value"), sep=";", row.names=T)
```



